---
output: 
  html_document: 
    keep_md: yes
---

title: "Reproducible Research Assessment 1"
output: html_document


#Downloading file and processing Data
-----------------------------------------------------------------


```{r}
if(!file.exists("getdata-projectfiles-UCI HAR Dataset.zip")) {
        activity<- tempfile()
        download.file("http://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip",activity)
        unzip(activity)
        unlink(activity)
}
activity<- read.csv("activity.csv", header = TRUE)
```

Let's check few Parameters

```{r}
head(activity)

summary(activity)
```

#What is mean total number of steps taken per day?
---------------------------------------------------
Calculate the total number of steps taken per day

```{r}
stepsperday<-aggregate(steps~date, activity, sum)
```

calculating Mean and Median of "steps per day"

```{r}
mean_step<-mean(stepsperday$steps)
mean_step

median_steps<-median(stepsperday$steps)
median_steps
```

Plotting Histogram
```{r}
hist(stepsperday$steps,col= "red", xlab="Total Number of Steps per day", ylab="Number of Days",
     main="Steps taken each day")
```

#What is the average daily activity pattern?
---------------------------------------------------
time series plot

```{r}
avgstepsbyinterval<-aggregate(steps~interval, activity, mean)
avgstepsbyinterval[which.max(avgstepsbyinterval[,2]),1]
with(avgstepsbyinterval, plot(interval, steps, type = "l",col= "blue"))
```

#Imputing missing values
----------------------------------------------------
Calculate and report the total number of missing values in the dataset

```{r}
missingvalues<-is.na(activity[,1])
```
replacing missing values with median

```{r}
med<-median(avgstepsbyinterval$steps)
activityNew<-activity
activityNew[missingvalues,1]<-med
head(activityNew)
```

Ploting histogram of the total number of steps taken each day 
after imputing missing values
 
```{r}
totalstepsperday1<-aggregate(steps~date, activityNew, sum)
hist(totalstepsperday1$steps,col = "blue", xlab="Total Steps per day", ylab="Days", 
     main="Number of Steps taken each day")
```

calculating the mean and median total number of steps per day 
we first find total number of steps per day

```{r}
totalstepsperday1<-aggregate(steps~date, activityNew, sum)

Newmean_stepsimput<-mean(totalstepsperday1$steps)

Newmean_stepsimput
```

Median

```{r}
med_stepsimput<-median(totalstepsperday1$steps)

med_stepsimput
```

#Are there differences in activity patterns between weekdays and weekends?
-----------------------------------------------------------------------------
```{r}
library(dplyr)
activityNew$date<-as.Date(activityNew$date)

activityNew2<-activityNew%>%
        mutate(dayType= ifelse(weekdays(activityNew$date)=="Saturday" 
                               | weekdays(activityNew$date)=="Sunday",
                               "Weekend", "Weekday"))
head(activityNew2)
```

Plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) 
and the average number of steps taken,
averaged across all weekday days or weekend days (y-axis)

```{r}
library(lattice)
averageStepsperDaytypeAndinterval<-activityNew2 %>%
        group_by(dayType, interval) %>%
        summarize(averageStepByDay=sum(steps))

head(averageStepsperDaytypeAndinterval)

with(averageStepsperDaytypeAndinterval, 
     xyplot(averageStepByDay ~ interval | dayType, 
            type = "l", col= "red",      
            main = "Total Number of Steps within Intervals by dayType",
            xlab = "Daily Intervals",
            ylab = "Average Number of Steps"))
```





